\chapter{Related Work}\label{chapter:related-work}
Cloud computing in the context of automotive allows for the realization of disruptive innovations which have great potential to improve driver satisfaction and to open up new fields of business for car manufacturers. For instance, \citeauthor*{wollaeger2012cloud} propose a framework which allows for the computation of an optimal velocity profile for vehicles in an effort to minimize fuel consumption \cite{wollaeger2012cloud}. Their solution benefits greatly from the vast amounts of computing resources available to clouds as it relies on dynamic programming which is characterized by high resource demands. Furthermore, large amounts of data are accessible in the cloud which may be used to make better-informed decisions, \eg\ on the basis of crowd-sourced data pools.

While some approaches exist which explore possible use cases of automotive clouds, very little work can be found that addresses the \emph{architectural} challenges of creating cloud-supported systems specifically for the automotive domain.
Thus, much of the work related to this thesis is concerned with other domains, such as IoT and fog computing. First and foremost, the thoroughly executed and exhaustive survey \cite{nath2018survey} on fog computing by \citeauthor*{nath2018survey} shall be mentioned as highly relevant and recommended reading. In their literature study, the authors present and summarize work pertaining to a variety of subjects related to IoT and embedded computing.
An important and much-discussed aspect in this domain is the \emph{management} of the of widely dispersed collections of embedded computing nodes that make up pervasive IoT systems. Similar to the system proposed in this thesis, many approaches exist that attempt to master the complexity inherent to these systems by means of service-oriented computing. Notable examples of service-oriented architecture (SOA) frameworks tailored specifically to the IoT and CPS are the SOCRADES project \cite{cannata2008socrades}, the IMC-AESOP project \cite{karnouskos2012soa} and the ARUM project \cite{marin2013conceptual}. Further noteworthy approaches are presented in \cite{butzin2016microservices}, \cite{kart2007distributed}, and \cite{teixeira2011service}. Of particular interest for this thesis are furthermore service-oriented approaches specifically tailored to \emph{automotive} systems. Contributions in this domain are provided, \eg , by \citeauthor*{kugele2017service}. In their paper \cite{kugele2017service}, the authors present the results of an interview study investigating the needs of automotive software architects and conclude that SOA-based approaches to E/E-system modeling bring promising benefits to the automotive development process.

Noteworthy in this regard is also SOME/IP,\footnote{``Scalable service-oriented MiddlewarE over IP''} an IP-based automotive middleware which may be used to model vehicular functionality in a service-oriented fashion, integrates flawlessly into AUTOSAR\footnote{``AUTomotive Open System ARchitecture''} and supports a large selection of communication paradigms.


\paragraph{Embedded SOA}

\begin{itemize}
%\item \cite {scholz2009soa}: \citeauthor*{scholz2009soa} - ∈ SOA-Service Oriented Architectures adapted for embedded networks
\item \cite{wagner2014developing}: \citeauthor*{wagner2014developing} - Developing self-adaptive automotive systems - On the integration of service-orientation into automotivedevelopment processes -- SOMA4DDAS
\end{itemize}


\paragraph{pubsub over SDN}
Of particular interest are approaches that aim to connect IoT devices with the cloud. Some approaches make use of software-defined networking (SDN) to achieve this. Notable work is presented, \eg \ by
\begin{itemize}
	\item \citeauthor*{farahzadi2017middleware} \cite{farahzadi2017middleware} analyze the problem domain of cloud connectivity for the IoT. For this purpose, they present a survey in which they analyze a number of different IoT middleware solutions and compile a set of key challenges and requirements for such systems.
	\item \citeauthor*{hakiri2015publish} \cite{hakiri2015publish} devise a system to interconnect IoT devices via DDS over an overlay network. However, the description of their approach lacks details and no real effort were made to evaluate it.
\end{itemize}


\paragraph{Containers + Embedded/IoT}
This thesis' approach relies on containerization as means to encapsulate functionality in isolated, portable units. A challenge in this is to marry containerization with the traditionally resource-constrained embedded systems. In recent years, this prospect has received substantial attention in the research community. For instance, \citeauthor*{morabito2017virtualization} provides an in-depth analysis on the utilization of container technology for a variety of Single-Board Computers (SBTs) \cite{morabito2017virtualization}. The author's evaluation includes benchmarks that assess energy consumption, device temperatures, as well as I/O and computing performance and other factors.
Further analyses and benchmarks are presented in \cite{pahl2015containers} and \cite{bellavista2017feasibility}.  
The evaluations reach a consensus that containerization for resource-constrained devices presents a promising prospect. 
In fact, the industry has already caught on.
resin.io\footnote{\url{www.resin.io}} is a project dedicated to bringing containers to embedded devices. In the past, resin.io have made several significant contributions to enable container-based Continous Integration for embedded systems, and have also taken initiative to further minimize resource requirements of containerization. To this end, resin.io provide their own light-weight container engine called ``balena''\footnote{\url{www.balena.io}} and an accompanying host OS that is deliberately crafted to run containers (``resinOS''\footnote{\url{www.resinos.io}}).


\paragraph{pubsub IoT}
Another building block of the approach presented in this thesis is the publish-subscribe communication paradigm.

\begin{itemize}
	\item \citeauthor*{antonic2016mobile} \cite{antonic2016mobile} present COPUS, a cloud-based publish-subscribe middleware for the IoT. In the paper, they perform benchmarks to evaluate the middleware with respect to scalability and elasticity and present a case study in which test subjects were equipped with mobile air quality monitoring sensors that would continuously send sensor data to the cloud. COPUS relies on a broker as cloud gateway which must be installed on a mobile device (\eg\ a smartphone). Thus, this solution is unsuitable for the automotive domain. 
\end{itemize}


\paragraph{Docker specific}
Relevant for the evaluation part of this thesis (\Cref{chapter:evaluation}) is \citeauthor*{kratzke2017microservices}'s paper \cite{kratzke2017microservices} on the performance impact of using encrypted overlay networks for Linux containers on top of hypervisors. This setup---containers atop hypervisors---is very common in cloud scenarios as IaaS providers often provision servers in the form of virtual machines. This is also the case for the test setup described in this thesis. The author concludes that containers add a non-negligible impact on networking performance. In \wnet -based overlay networks, the performance is further impaired, and another, less-significant impact is observed when employing encryption. This is in contrast to the benchmark results presented in this thesis, which suggest that both technologies incur a very minor overhead.

% does not consider weave fast datapath
% argues that weave would be faster if the weave routers weren't packaged in containers



\paragraph{DDS + Virtualization}
\begin{itemize}
	\item \cite{serrano2013virtualizing}: Virtualizing DDS middleware: performance challenges and measurements
	\item \citeauthor*{perez:gutierrez:ieeetpds16} \cite{perez:gutierrez:ieeetpds16} present a way to combine DDS and hypervisor-based inter-process communication to interconnect a time and space partitioned system. Their focus lies on mixed-criticality applications running atop a hypervisor designed for safety-critical scenarios. The performance tests they conduct show that DDS adds a reasonable overhead to the communication performance but the overhead caused by hypervisor-based partitioning is quite significant, albeit reducible by the use of multicore processors.
	\item \cite{garcia2013benchmarking}: Benchmarking communication middleware for cloud computing virtualizers. Test performance overhead of hypervisors for DDS as well as the performance overhead of DDS compared to ICMP packets
\end{itemize}


\paragraph{Publish-Subscribe in Embedded Systems}
\begin{itemize}
	\item \cite{bouhouch2013dds}: DDS on top of FlexRay
	\item \cite{sommer2013race}: RACE: A centralized Platform computer based architecture for automotive applications
	\item \cite{gundougan2018hopp}: HoPP: Robust and Resilient Publish-Subscribe for an Information-Centric Internet of Things
	\item SOME/IP
\end{itemize}


\paragraph{Cloud + IoT}

\begin{itemize}
	\item \cite{nastic2014provisioning}: Provisioning Software-defined IoT Cloud Systems: suggest system for software-defined IoT. Encapsulate IoT devices as software artifacts that can be controlled via API. This allows a great deal of automation and run-time customizations
\end{itemize}


\paragraph{Cloud}

\begin{itemize}
	\item \cite{greenberg2008cost}: The cost of a cloud: research problems in data center networks. Insights: location-independent addressing is a major design objective
	\item \cite{wang2010impact}: The impact of virtualiztion on network performance. The authors perform networking benchmarks on Amazon EC2 instances and find that virtualization and processor sharing heavily impacts networking performance. In particular, network delay and TCP/UDP throughput suffers drastically.
\end{itemize}

\paragraph{computation offloading in cloud (cell phones)}
\citeauthor*{barbera2013offload} \cite{barbera2013offload} analyze in which cases it is sensible to offload computations from cell phones to the cloud. There is a tradeoff to be made between the energy savings achieved by the offloading itself and the increased computational overhead of sending data over a mobile network. They aim to provide virtual replicas of the hardware in the cloud, running at all times alongside the physical device to support it whenever there is a need. They do not consider statelessness of the replicas and employ a synchronization mechanism to keep the state consistent. This incurs additional overhead. They use cell phone specific tools which are inapplicable in the automotive use case. They view their approach as holistic solution -> lack of flexibility

In \cite{chun2011clonecloud} \citeauthor*{barbera2013offload} present \emph{Clone Cloud}, a system to offload workload from mobile applications to the cloud. They employ a combination of static analysis and dynamic profiling to automatically partition mobile applications. repartitioning is thread-wise. They achieve a 20-fold execution speed-up and 20-fold decrease of energy consumption.

MAUI is not as automatic. Requires program code to be annotated.

\paragraph{Other}
\citeauthor*{broy2006challenges} \cite{broy2006challenges}: Challenges in automotive software engineering... Takeaways: increasing amounts of code in cars, increasing complexity. Software is the main innovation driver. Broy states challenges, gives an industry profile of automotive industry. Lists requirements that are very different from other domains. Current development practices in the industry are lacking.





\paragraph{Similar Approaches}
Finally, after having covered the basic building blocks of this thesis' approach, there is some related work that takes similar measures to achieve similar goals.
One example is presented in \cite{berger2017containerized}. In their paper, \citeauthor*{berger2017containerized} describe their experiences of employing containerization in the context of self-driving vehicles.
In accordance with the microservice architecture pattern, vehicular functionality in their approach is split into a number of fine-grained services deployed on distributed nodes within the vehicle. The services are connected via OpenDaVINCI,\footnote{\url{code.opendavinci.org}} a real-time-capable middleware for distributed embedded systems. This approach is very similar to the one proposed in this thesis in that communication is based on multicast-enabled publish-subscribe and that they employ \docker\ containers. However, their approach lacks cloud connectivity. The authors furthermore struggle to get multicast to work over Docker overlay networks---a problem that was solved in this thesis.
%The interplay of their services is based on a \emph{pipes-and-filters} methodology.
%Multi platform images are achieved through different base layers employing different compilers. Disadvantage: containers cannot be moved freely between nodes of different hardware architectures.
%questionable design decisions: Centralized configuration protocol... but decentralization is key
%Suggest to have private registry within the vehicle.
%To provide a high degree of tracability, they present a versioning system that allows versioning of individual image layers which seamlessly integrates into their development workflow.
%Similarly, \citeauthor*{schneider2016achieving} \cite{schneider2016achieving} use containerization in a microservice environment for vehicular functionality. However, the authors only focus on the backend part, \ie , services provisioned in data centers that vehicles may connect to. Their approach does not consider the use of microservices \emph{within} the vehicle. In addition, their methodology is described in rather vague terms and lacks detail.

Another approach that resembles the one presented in this thesis is described by Asad \citeauthor*{javed2016container}. In his master's thesis \cite{javed2016container}, he explores the possibility of managing IoT nodes via \docker\ containers and the Kubernetes\footnote{\url{www.kubernetes.io}} container orchestration platform. The employed experimental setup is very similar to the one used in this thesis in that a local cluster of Raspberry Pis is connected in a LAN with cloud access. Attached to some of the Raspberry Pis are cameras that capture images which are then sent to the cloud via Apache Kafka. Like in this thesis, communication between the nodes and the cloud is publish-subscribe-based. However, the approach has a serious drawback: Kubernetes and Kafka are technologies designed for sophisticated cloud environments and are by no means suitable for the use within sensor networks due to their tremendously high system requirements. Moreover, no serious effort is evident to evaluate the approach in a neutral manner.
\citeauthor*{grossmann2016hypriot} \cite{grossmann2016hypriot} describe another approach related to the previous one. But instead of heavy-weight enterprise container management solutions they follow a more light-weight philosophy. A drawback of their approach is a lack of failure resiliency inherent to their centralized architecture in which a master node may become a single point of failure. A follow-up to that work is presented by \citeauthor*{celesti2017watchdog} \cite{celesti2017watchdog} in which several weak points are addressed.

%The fact that ubiquitous publish-subscribe is not only a research matter but is also the basis of multiple commercial industry products is proof for its value. 
Not only in the research community, but also in the industry, there is interest in ubiquitous publish-subscribe. 
For instance, PubNub\footnote{\url{www.pubnub.com}} is a commercial service that aims to provide publish-subscribe-style communication for large numbers of devices dispersed on a global scale. Their focus lies on IoT devices and mobile phone applications. Thus, there is a pronounced focus on handling mobility challenges such as frequently changing IP addresses and general reliability issues. PubNub is a holistic, all-in-one solution, \ie , their service includes everything from client APIs to infrastructure provisioning. The fact that \mbox{PubNub} is a closed platform severely inhibits its adaptability and limits its applicability---especially for the automotive use case, which is characterized by highly specific requirements. While it may be true that PubNub is a good starting point to kick off new businesses (as advertised), it is doubtful that it is the right tool for the intended use case.

Another commercial product that follows the same goal, but is focused more on Industrial IoT (IIoT), is offered by ADLINK (formerly PrismTech). ADLINK provides a holistic data sharing platform based on DDS named ``Vortex DDS''. Aside from their own DDS implementation, OpenSplice, the Vortex platform includes the service called ``Vortex Cloud''.\footnote{\url{www.prismtech.com/vortex/vortex-cloud}} Like this thesis' approach, Vortex Cloud aims to connect globally dispersed DDS components, and also provides means to deploy DDS-based applications in the cloud. Unfortunately, no information about the inner workings of Vortex Cloud and its aptitude can be found online. As it is a commercial, closed-source product, this option was not further pursued.

Real-Time Innovations (RTI) is another company which provides an DDS implementation called ``RTI Connext DDS''.\footnote{\url{www.rti.com/products}} A salient feature of Connext is that it supports DDS-based data transmission over WANs natively. However, this ``RTI Secure WAN Transport'' has several drawbacks. Firstly, it needs to be configured on code-level in all applications that participate in the communication network. This is in contrast to this thesis' approach in which the applications themselves are kept unaware of the fact that they are communicating with peers outside their LAN. This agnostic approach promotes loose coupling, simplifies network management, and ultimately enables applications to migrate freely between computing nodes. Further disadvantages of RTI's approach are that it doesn't support multicast, that it requires a central ``rendezvous'' server which acts as a participant repository, and that it is not resilient to connection loss.\footnote{\url{community.rti.com/static/documentation/connext-dds/5.3.0/doc/manuals/connext_dds/html_files/RTI_ConnextDDS_CoreLibraries_UsersManual/index.htm}}

%For this, a central, publicly accessible ``Cloud Discovery Service'' is required which acts as a repository that holds all participating applications' locations.


